---
title: "ESM 204 Assignment 4"
author: "Anna Talken, Sydney Rilum & Grace Kumaishi"
date: "5/18/2021"
output: 
  html_document:
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
```

```{r}
# Read in data:
damages <- read.csv(here("damages.csv"))
warming <- read.csv(here("warming.csv"))
```

## Question 1:

### Estimate the quadratic damage function:
```{r}
# Visualize points:
# ggplot(data = damages, aes(x = warming, y = damages)) +
  # geom_point()

# Create a new data frame for warming^2:
damages$warming2 <- damages$warming^2 

# Fit quadratic regression model:
quadratic.model <-lm(damages ~ warming + warming2 + 0, data = damages)

# View model summary:
summary(quadratic.model)
```

#### Estimated damages quadratic model:
$$y = (1.94e13) x^2 - (1.03e12)x$$

```{r fig.align = "center"}
### Plot estimated damage:
x = damages$warming
a = 1.959e+13
b = -1.03e+12
y = a * x^2 + b * x

damages_function <- plot(x,y, type = "l")

ggplot(data = damages, aes(x = warming, y = damages)) +
  geom_point() +
  geom_smooth(formula = damages_function, color = "darkred")+
  theme_minimal()+
  labs(title = "Predicted damages based on increasing global mean temperatures",
       x = "Warming (degrees C)",
       y = "Damages ($)") +
  theme(plot.title = element_text(hjust = 0.5))
```


## Question 2

### 1) Plot damages over time without pulse:

```{r fig.align = "center"}
# Create new function to calculate damages based on damages quadratic equation (didn't actually need this....):
damages <- function(w, model){
  d <- model$coefficients[[3]]*w^2 + model$coefficients[[2]]*w
  return(d)
}

# Cost of damages based on baseline climate scenario:

warming <- warming %>% 
  mutate(damages_baseline = a*warming_baseline^2 + b*warming_baseline,
         damages_pulse = a*warming_pulse^2 + b*warming_pulse,
         dif_pulse = damages_pulse - damages_baseline,
         dif_pulse_perton = dif_pulse/35000000000)

ggplot(data = warming, aes(x = warming_baseline, y = damages_baseline)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Future damages based on baseline warming estimates",
       x = "Baseline warming (degrees C)",
       y = "Damages ($)") +
  theme(plot.title = element_text(hjust = 0.5))
```

### 2) Plot damages over time with the pulse:

```{r fig.align = "center"}
ggplot(data = warming, aes(x = warming_pulse, y = damages_pulse)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Future damages based on warming estimates with a pulse",
       x = "Warming with pulse (degrees C)",
       y = "Damages ($)") +
  theme(plot.title = element_text(hjust = 0.5))
```

### 3) Difference in damages over time that arise from pulse:

```{r fig.align = "center"}
ggplot(data = warming, aes(x = X, y = dif_pulse)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Differences in future damages with a pulse",
       x = "Year",
       y = "Difference in damages ($)") +
  theme(plot.title = element_text(hjust = 0.5))
```

### 4) Difference in damages over time from the pulse per ton of CO2:

```{r fig.align = "center"}
ggplot(data = warming, aes(x = X, y = dif_pulse_perton)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Differences in future damages with a pulse per ton",
       x = "Year",
       y = "Difference in damages per ton ($)") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Question 3:

#### Use the equation: `V0 = Vt / (1 + r)^t`

```{r fig.align = "center"}
# Create vector with 5 different discount rates:
rate <- c(0.01, 0.02, 0.03, 0.04, 0.05)

# Vt/(1+r)^t

SCC <- c()
for (i in 1:length(rate)) {
    cost <- (warming$dif_pulse_perton)/(1 + rate[i])^(warming$X)
    SCC[i] <- sum(cost)
  }

# Create table with SCC for each discount rate:
SCC_table <- round(data.frame("discount_rates" = rate,
                        "SCC" = SCC), 2)

# Graph discount rate vs. SCC:
ggplot(data = SCC_table, aes(x = rate,
                             y = SCC)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Social Cost of Carbon (SCC) at varied discount rates",
       x = "Discount rate",
       y = "SCC") +
  theme(plot.title = element_text(hjust = 0.5))

```

## Question 4:

```{r fig.align = "center"}
# r = p +ng
ramsey = 0.021

# Create vector with 5 different discount rates:
ramsey_rate <- c(0.01, 0.02, 0.03, 0.04, 0.05, ramsey)

# Vt/(1+r)^t

SCC_ramsey <- c()
for (i in 1:length(ramsey_rate)) {
    cost_ramsey <- (warming$dif_pulse_perton)/(1 + ramsey_rate[i])^(warming$X)
    SCC_ramsey[i] <- sum(cost_ramsey)
}

SCC_table_ramsey <- data.frame("discount_rates" = ramsey_rate,
                        "SCC" = SCC_ramsey)

### Based on the output from the table above, we know that the new SCC value should be 71.35 ###

# Graph ramsey discount rate vs. SCC:
ggplot(data = SCC_table, aes(x = rate,
                             y = SCC)) +
  geom_point() +
  geom_line() +
  geom_point(x = 0.021, y = 71.35061, color = "red", size = 3) +
  theme_minimal() +
  labs(title = "Ramsey point overlayed with SCC/discount rate graph",
       x = "Discount rate",
       y = "SCC") +
  theme(plot.title = element_text(hjust = 0.5))

```


## Question 5:

### What is the expected present value of damages up to 2100 under Policy A?

```{r}

```

### What is the expected present value of damages up to 2100 under Policy B?

```{r}

```

### If Policy A costs $0 and Policy B costs $X, how large could X be for it to still make economic sense to pursue Policy B instead of A?

```{r}

```

### How would your answer change if society were risk averse? (Qualitative answer)
